---
title: "bomWater"
author: "Alexander Buzacott"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{bomWater}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(httptest)
start_vignette("bomWater")
```

## Introduction

`bomWater` provides an interface to the Australian Bureau of Meteorology Water Data online (http://www.bom.gov.au/waterdata/) web service. As part of the Water Act 2007, the BoM is required to collect Australian water information and make it available to the public. This provides a single and convenient source of water data for all states and territories. Several hydrological and meteorological quality checked timeseries are available through Water Data online and `bomWater` provides several functions to access these timeseries.

## Using bomWater

A typical workflow begins by querying station information using `get_station_list()` with a specific parameter type. The function `parameters()` returns all the parameters that can be retrieved using `bomWater`. By default `get_station_list()` returns all stations for Water Course Discharge and other parameters can be specified with the argument `parameter_type`. Once the station number of interest is known, it can be used to retrieve quality-checked timeseries using the functions:

* `get_hourly()`: retrieves hourly data
* `get_daily()`: retrieves daily data
* `get_monthly()`: retrieves monthly data
* `get_yearly()`: retrieves annual data
* `get_as_stored()`: retrieves data as stored by BoM

Timeseries that are based on aggregated data return the mean by default, however other variables such as min and max are often available. The functions will return a `tibble` with three columns: Timestamp, Value and Quality Code.

### Parameters

The following Water Data Online variables can be accessed using these functions:

| Parameter                      | Units  |
| ------------------------------ | ------ |
| Water Course Discharge         | m3/s   |
| Water Course Level             | m      |
| Electrical conductivity at 25C | µS/cm  |
| Turbidity                      | NTU    |
| pH                             | pH     |
| Water Temperature              | ºC     |
| Storage Volume                 | ML     |
| Storage Level                  | m      |
| Ground Water Level             | m      |
| Rainfall                       | mm     |
| Evaporation                    | mm     |
| Dry Air Temperature            | ºC     |
| Relative Humidity              | %      |
| Wind Speed                     | m/s    |

### Quality codes

The following table is from the ([BoM SOS2 manual](http://www.bom.gov.au/waterdata/wiski-web-public/Guide\%20to\%20Sensor\%20Observation\%20Services\%20(SOS2)\%20for\%20Water\%20Data\%20\%20Online\%20v1.0.1.pdf)) (6.3.3, pg 41) and summarises the quality codes.

| SOS2 Qualifier | Quality code | Description |
| -------------- | ------------ | ----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------- |
| 10  | A | The record set is the best available given the technologies, techniques and monitoring objectives at the time of classification. |
| 90  | B | The record set is compromised in its ability to truly represent the parameter. |
| 110 | C | The record set is an estimate. |
| 140 | E | The record set's ability to truly represent the monitored parameter is not known. |
| 210 | F | The record set is not of release quality or contains missing data. |

## Load packages

Load `bomWater` as well as `ggplot` to make some plots.

```{r setup}
library(bomWater)
library(ggplot2)
```

## Using parameters()

The parameter names can be retrieved from `bomWater` using `parameters()`. `parameters()` takes no argument. Make sure the formatting of the parameter types is as returned by this function otherwise an error may be raised.

### Results

Returns a vector of parameter types.

### Example

```{r}
parameters()
```

## Using get_station_list()

'get_station_list' queries Water Data Online and returns station details. Queries can be input with the desired 'parameter_type' to find all the stations on record. If you already have a vector of station numbers, you can pass the vector to 'station_number' and return the details of those stations. 'return_fields' can be customised to return various data about the stations.

### Results

With the default return fields, a tibble with columns station_name, station_no, station_id, station_latitude, station_longitude.

### Example

```{r}
# Get a list of groundwater bores available from water data online
get_station_list(parameter_type = "Ground Water Level")

# Return information for a single station and customise return fields
get_station_list(station_number = "410730")
```

## Using get_hourly()

`get_hourly` returns hourly data. Only Water Course Discharge, Water Course Level, Storage Level and Storage Volume can be requested using `get_hourly`.

### Result

Returns a tibble with Timestamp, Value and Quality Code.

### Example

```{r}
# Cotter River at Gingera
get_hourly(
  parameter_type = "Water Course Discharge",
  station_number = "410730",
  start_date = "2020-01-01",
  end_date = "2020-01-31"
)

# Corin Reservoir
get_hourly(
  parameter_type = "Storage Volume",
  station_number = "410742",
  start_date = "2020-01-01",
  end_date = "2020-01-31"
)
```

An example of plotting the data:

```{r}
corin <- get_hourly(
  parameter_type = "Storage Volume",
  station_number = "410742",
  start_date = "2020-01-01",
  end_date = "2020-01-31"
)
ggplot(corin, aes(Timestamp, Value)) +
  geom_line() +
  labs(x = "Time", y = "Storage Volume (ML)")
```

## Using get_daily()

Retrieves timeseries at a daily timestep. For continuous data, the daily mean (default), max, and min can be returned by specifying the `var` argument. Daily totals are returned for discrete parameters (e.g. rainfall and evaporation). Timeseries aggregated between 9am to 9am for Rainfall and Water Course Discharge can be retrieved using the `aggregation` argument to match the standard BoM rainfall reporting.

### Result

A tibble with columns Timestamp, Value and Quality Code.

### Examples

```{r}
# Daily mean streamflow from Cotter River at Gingera (in m3/s) between 09-09
get_daily(
  parameter_type = "Water Course Discharge",
  station_number = "410730",
  start_date     = "2020-01-01",
  end_date       = "2020-01-31",
  aggregation    = "09HR"
)

# Daily max, only available over the standard day
get_daily(
  parameter_type = "Water Course Discharge",
  station_number = "410730",
  start_date     = "2020-01-01",
  end_date       = "2020-01-31",
  var            = "max"
)

# Daily mean wind speed at Corin Dam
get_daily(
  parameter_type = "Wind Speed",
  station_number = "570947",
  start_date     = "2020-01-01",
  end_date       = "2020-01-31"
)
```

## Using get_monthly()

Retrieves timeseries at a monthly timestep. Monthly totals are returned for discrete parameters (e.g. rainfall and evaporation), while the mean rate is returned for continuous parameters.

### Result

A tibble with columns Timestamp, Value and Quality Code.

### Examples

```{r}
# Monthly total rainfall in mm at Cotter Hut
get_monthly(
  parameter_type = "Rainfall",
  station_number = "570946",
  start_date     = "2019-01-01",
  end_date       = "2019-12-31"
)

# Monthly mean streamflow rate m3/s at Cotter River at Gingera
get_monthly(
  parameter_type = "Water Course Discharge",
  station_number = "410730",
  start_date     = "2019-01-01",
  end_date       = "2019-12-31"
)

# Monthly evaporation at Blowering Dam
get_monthly(
  parameter_type = "Evaporation",
  station_number = "410102",
  start_date     = "2019-01-01",
  end_date       = "2019-12-31"
)
# No data
```

## Using get_yearly()

Retrieves timeseries at an annual timestep. Annual totals are returned for discrete parameters (e.g. rainfall and evaporation), while the mean rate is returned for continuous parameters.

### Result

A tibble with columns Timestamp, Value and Quality Code.

### Examples

```{r}
# Annual rainfall at Berthong in Cootamundra
berthong <- get_yearly(
  parameter_type = "Rainfall",
  station_number = "41000207",
  start_date     = "2010-01-01",
  end_date       = "2019-12-31"
)

# Example plot
ggplot(berthong, aes(Timestamp, Value)) +
  geom_col() +
  labs(x = "Time", y = "Rainfall (mm/year)")
```

## Using get_as_stored()

Retrieves timeseries as stored by the BoM. Stored frequency could vary substantially between different parameters and locations.

### Result

A tibble with columns Timestamp, Value and Quality Code.

### Examples

```{r}
# Annual rainfall at Berthong in Cootamundra
berthong <- get_as_stored(
  parameter_type = "Rainfall",
  station_number = "41000207",
  start_date     = "2019-01-01",
  end_date       = "2019-12-31"
)
berthong
```

## Copyright

The license and copyright for the data can be viewed under the copyright tab at http://www.bom.gov.au/waterdata/. More information about the Bureau of Meteorology copyright can be found at http://www.bom.gov.au/other/copyright.shtml.

```{r, include=FALSE}
end_vignette()
```
